---
title: "Correlation using R"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Loading Data

Let us load the data for the further processing.

```{r}
library(RCurl)

urlfile <- 'https://archive.ics.uci.edu/ml/machine-learning-databases/adult/adult.data'
x <- getURL(urlfile, ssl.verifypeer = FALSE)
adults <- read.csv(textConnection(x), header=F)

names(adults)

names(adults)<- c('Age','Workclass','FinalWeight','Education','EducationNumber',
                'MaritalStatus','Occupation','Relationship','Race',
                'Sex','CapitalGain','CapitalLoss','HoursWeek',
                'NativeCountry','Income')

adults$Income <- ifelse(adults$Income==' <=50K',1,0)
```

## Convert Categorical Variable into Numerical

The correlation work for numerical values only. So first we have to convert the categorical variable into numerical variable using powerful package **caret**. In this package we use **dummyVars** funcion for coverting the categorical variable.


```{r}
library(caret)
dmy <- dummyVars(" ~ .", data = adults)
adultsTrsf <- data.frame(predict(dmy, newdata = adults))
```

## Correlation matrix with p-values

Then define **cor.prob** which give the correlation matrix with p-values.

```{r}
cor.prob <- function (X, dfr = nrow(X) - 2) {
  R <- cor(X, use="pairwise.complete.obs")
  above <- row(R) < col(R)
  r2 <- R[above]^2
  Fstat <- r2 * dfr/(1 - r2)
  R[above] <- 1 - pf(Fstat, 1, dfr)
  R[row(R) == col(R)] <- NA
  R
}
```

Use this cor.prob output to a 4 column matrix with row/column indices, correlation, and p-value.

```{r}
flattenSquareMatrix <- function(m) {
  if( (class(m) != "matrix") | (nrow(m) != ncol(m))) stop("Must be a square matrix.")
  if(!identical(rownames(m), colnames(m))) stop("Row and column names must be equal.")
  ut <- upper.tri(m)
  data.frame(i = rownames(m)[row(m)[ut]],
             j = rownames(m)[col(m)[ut]],
             cor=t(m)[ut],
             p=m[ut])
}

corMasterList <- flattenSquareMatrix (cor.prob(adultsTrsf))
print(head(corMasterList,10))
```

## Choose variable of high correlation

Sort the matrix by correlation value 

```{r}
corList <- corMasterList[order(corMasterList$cor),]
```

Subset those variable which is highly correlated with target variable.

```{r}
selectedSub <- subset(corList, (abs(cor) > 0.2 & j == 'Income'))

bestSub <-  sapply(strsplit(as.character(selectedSub$i),'[.]'), "[", 1)
```

Plot the selected variable

```{r}
library(psych)
pairs.panels(adults[c(bestSub,'Income')])
```


